Global deaths due to air ollution

### 2022-04-30 outline.Rmd
re-arranged chunks in outline

---
title: "Global deaths due to air ollution"
author: "Elizabeth Bekele, Alison Cheek"
output: slidy_presentation
date: '2022-05-03'
---

# Introduction 
* Air pollution can be detrimental to both our health and the climate
  + Outdoor and indoor air pollution cause chronic pain, respiratory diseases, shortened lifespan
  + Air pollution kills about 7 million people worldwide every year 
  + Hopefully this information can showcase the importance of air pollution and that we should be more mindful about our planet

@@

# Data Details 
*Air pollution deaths 
  +Kaggle: https://www.kaggle.com/datasets/akshat0giri/death-due-to-air-pollution-19902017 
  + Author: Akshat Giri
  + Last update: 2 years ago

```{r, include=FALSE}
#This will allow us to filter through our data 
library(tidyverse)
library(dplyr)
#This will help us plot figures to showcase our findings
library(ggplot2)
#This will help us organize and display our data as necessary 
library(knitr)
library(kableExtra)
#This expands our plot uses 
library(plotly)
#Adjusting numbers 
format(4.470862e-01, scientific = F, digits = 3)
```

# Data Preparation 
The data we will be looking at:

1. **Deaths due to Air Pollution**
[https://www.kaggle.com/datasets/akshat0giri/death-due-to-air-pollution-19902017]

* Author: Akshat Giri
* This data consists the deaths due to air pollution from 1990-2017 around the world
* The dataset was recently updated 2 years ago
* The author wanted to gain more awareness about air pollution and forecast the future

2. **World Population(pending)**

First let's import the Deaths by Air Pollution data.
```{r}
deaths_df <- data.frame(read.csv("death-rates-from-air-pollution.csv"))
```

@@

Now, we are going to import the World Population data 
```{r}
world_pop <- read.csv("population_total_long.csv")
glimpse(world_pop)
```

Let's split the data into high and low population based on country.
- Low population = high population * .10
```{r}
#Let's compare the first decade 1996-2006 with second decade 2007-2017 
high_pop_countries <- world_pop %>% 
  filter(Year > 1996 & Country.Name %in% c('United States', 'Brazil', 'Nigeria', 'Germany', 'Pakistan', 'Australia')) %>% 
  group_by(Year)

@@

low_pop_countries <- world_pop %>% 
  filter(Year > 1996 & Country.Name %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'New Zealand'))%>% 
  group_by(Year)

@@

```
After selecting our desired years, we've cut the population data set nearly in half.

To get a general idea of the data, let's make some plots to see what's happening.

```{r}
d <- ggplot(deaths_df, aes(x = indoor_deaths, y = outdoor_deaths, text = paste0(country, ", ", year) )) + geom_point()
ggplotly(d)

#p <- ggplot(world_pop, aes(x = Country.Name, y = Year,  text = paste0(Country.Name, ", ", Year) )) + geom_density()
#ggplotly(p)
```

>>>>

---
title: "Global deaths due to air ollution"
author: "Elizabeth Bekele, Alison Cheek"
output: slidy_presentation
date: '2022-05-03'
---
```{r, include=FALSE}
#This will allow us to filter through our data 
library(tidyverse)
library(dplyr)
#This will help us plot figures to showcase our findings
library(ggplot2)
#This will help us organize and display our data as necessary 
library(knitr)
library(kableExtra)
#This expands our plot uses 
library(plotly)
```
# Introduction 
* Air pollution can be detrimental to both our health and the climate
  + Outdoor and indoor air pollution cause chronic pain, respiratory diseases, shortened lifespan
  + Air pollution kills about 7 million people worldwide every year 
  + Hopefully this information can showcase the importance of air pollution and that we should be more mindful about our planet

@@

# Data Details 
*Air pollution deaths 
  +Kaggle: [https://www.kaggle.com/datasets/akshat0giri/death-due-to-air-pollution-19902017] 
  + Author: Akshat Giri
  + Last update: 2 years ago

Import the Deaths by Air Pollution data
```{r}
deaths_df <- data.frame(read.csv("death-rates-from-air-pollution.csv"))
```

@@

Now, we are going to import the World Population data 
```{r}
world_pop <- read.csv("population_total_long.csv")
head(world_pop)
```
To get a general idea of the data, let's make some plots to see what's happening.

```{r}
d <- ggplot(deaths_df, aes(x = indoor_deaths, y = outdoor_deaths, text = paste0(country, ", ", year) )) + geom_point()
ggplotly(d)
```

Let's split the data into high and low population based on country

  +Low population = high population * .10

```{r}
#selecting high-population countries from the world population data frame 
high_pop_countries <- world_pop %>% 
  filter(Year > 1996 & Country.Name %in% c('United States', 'Brazil', 'Nigeria', 'Germany', 'Pakistan', 'Australia')) %>% 
  group_by(Year)

@@

#selecting low-population countries from the world population data frame 
low_pop_countries <- world_pop %>% 
  filter(Year > 1996 & Country.Name %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'New Zealand'))%>% 
  group_by(Year)

@@
